Verifying a Postfix Language \\ \large CS595 Final Project Proposal

### 2021-11-25 project/paper.tex
start on final paper

% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

@@

\begin{document}
\maketitle
\begin{abstract}
We intend to formalize and give some proofs of soundness to a subset of \href{https://github.com/dvtate/postfix-haskell}{a functional, postfix language}.
\end{abstract}

@@

\section{Syntax}
Taking the ideas from the above rant into account we arrive at something along these lines
\[
\begin{array}{r l}
    \tau \bnfdef & \kwInt \bnfalt \kwLong \bnfalt \kwFloat \bnfalt \kwDouble \bnfalt 
        (\tau^*) \bnfalt \kwUnit \bnfalt \kwVoid \bnfalt \tau_1~ \tau_2~ + 
        \bnfalt ~ \kwArrow{\tau_1} {\tau_2} \bnfalt \kwType{e}\\

@@

    S \bnfdef & S~ S \bnfalt e~ \kweid{x} = \bnfalt ~e

@@

\section{Proposed Work}
\subsection{The Formalization}
Using big step semantics most of these aren't too scary. Just need to sit down and combine ideas.

\subsection{Proof of Soundness}
Proving theorems of Progress and Preservation should sufficiently prove a baseline of safety.

\subsection{Findings, Resolution?, Conclusions}
It's entirely possible that we could encounter problems with the language design, although pretty unlikely since we're even requiring type signatures on closures.

\section{Plan Distribution of Work}
"Mostly" is used to indicate any >50\% but likely not independent
\subsection{Timeline}
\begin{itemize}
    \item Week of Nov 15: Further research and discussion over formalization, rules written down, mostly Tate
    \item Week of Nov 22: Tate and Andrew will each independently work on the soundness proofs (Tate Progress, Andrew preservation, although subject to change)
    \item Week of Nov 29: 
        \begin{itemize}
            \item Peer-review proofs - equal work
            \item Combine proofs into paper - mostly Andrew
            \item Start on presentation - mostly Andrew
        \end{itemize}
    \item Week of Dec 6: 
        \begin{itemize} 
            \item Attempt to resolve any findings - mostly Tate
            \item finalize paper presentation - equal work
        \end{itemize}
\end{itemize}

\subsection{Things to Research}
\begin{itemize}
    \item Formalization for Forth or PostScript, stack-based languages which don't follow the variable binding
\end{itemize}

>>>>

% Useful packages
\usepackage{amsmath, amsthm, verbatim, enumerate, xcolor, mathpartir}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

@@

\begin{document}
\maketitle
\begin{abstract}
A formalization and some proofs of soundness is given to a subset of \href{https://github.com/dvtate/postfix-haskell}{a functional, postfix language}.
\end{abstract}

@@

\section{Syntax}
Taking the ideas from the above rant into account we arrive at something along these lines
\[
\begin{array}{r l}
    \tau \bnfdef & \kwInt \bnfalt \kwLong \bnfalt \kwFloat \bnfalt \kwDouble \bnfalt 
        (\tau^*) \bnfalt \kwUnit \bnfalt \kwVoid \bnfalt \tau_1~ \tau_2~ \texttt{|} 
        \bnfalt ~ \kwArrow{\tau_1} {\tau_2} \bnfalt \kwType{e}\\

@@

    S \bnfdef & \varepsilon \bnfalt S~ S \bnfalt e~ \kweid{x} = \bnfalt ~e

@@

\section{Semantics}
We've expanded the macro value syntax to also capture lexically scoped variables. These are represented as with a brackets syntax as shown: $(( \tau^* )[ x \step (), y \step 123 ]: ... )$.

{
    \centering
    \def \MathparLineskip {\lineskip=0.43cm}
    \begin{mathpar}
        \Rule{Num Lits}
             {\forall ~ T \in \{ \kwInt, \kwLong, \kwFloat, \kwDouble\}}
             {\kwnumlit{n}{T} \val}
        \and

        % Derivation of e* -> e*'
        \Rule{$e^* \step~ e^{*\prime}$ - Part 1}
            {e_1 \step e_1'}
            {e_1 ~ ... ~ e_n \step e_1' ~ ... ~ e_n}
        \and
        \Rule{$e^* \step~ e^{*\prime}$ - Part 2}
            {e_1 \step e_1'
                \\v_i \val ~ \forall v_i \in v*}
            {v* ~ e_1 ~ ... ~ e_n \step v* ~ e_1' ~ ... ~ e_n}
        \and

        % Tuple literals evaluation
        \Rule{Tuple Lits - 1}
            {e^* \step e^{*\prime}}
            {( e^* ) \step ( e^{*\prime} )}
        \Rule{Tuple Lits - 2}
            {v_i \val ~ \forall v_i \in v*}
            {(v*) \val}
        \and

        % unpack operator
        \Rule{Tuple Value Unpack - 1}
            {v_i \val ~ \forall i \in [1, n]}
            {\kwunpack{(v_1 ... v_n)} \step v_1 ... v_n}
        \Rule{Tuple Value Unpack - 2}
            {\strut}
            {\kwunpack{()} \step \varepsilon}
        \and
        \Rule{Tuple Type Unpack - 1}
            {\strut}
            {\kwunpack{(\tau_1 ... \tau_n)} \step \tau_1 ... \tau_n}
        \Rule{Tuple Type Unpack - 2}
            {\strut}
            {\kwunpack{\kwUnit} \step \varepsilon}

        % math
        \Rule{Numeric Sum}
            {v_1 \val \\ v_2 \val \\ v_3 = v_1 + v_2}
            {v_1~ v_2~ + \step v_3}
        \Rule{Numeric Product}
            {v_1 \val \\ v_2 \val \\ v_3 = v_1 * v_2}
            {v_1~ v_2~ \times \step v_3}
        \and

        % ternary
        \Rule{Ternary 1}
            {v_1 \val \\ v_2 \val}
            {\kwtern{\kwnumlit{0}{\kwInt}}{v_1}{v_2} \step v_2}
        \Rule{Ternary 2}
            {v_1 \val \\ v_2 \val \\ n \neq 0}
            {\kwtern{\kwnumlit{n}{\kwInt}}{v_1}{v_2} \step v_1}

    \end{mathpar}
}
